\documentclass[11pt,a4paper]{article}

% ===== Packages =====
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{mathtools}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  citecolor=green!50!black,
  urlcolor=blue!50!black
}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  framerule=0.3pt,
  framesep=4pt,
  columns=fullflexible,
  showstringspaces=false
}
\usepackage{algorithm}
\usepackage{algpseudocode}

% ===== Theorem environments =====
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\theoremstyle{plain}
\newtheorem*{conjecture*}{Conjecture}
\newtheorem{conjecture}[theorem]{Conjecture}

% ===== Macros =====
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\ZZp}[1]{\mathbb{Z}_{#1}}
\newcommand{\Branch}{\mathcal{B}}
\newcommand{\InvG}{G^{\leftarrow}}
\newcommand{\Calphax}{\mathcal{C}^{\leftarrow}_{S202}}
\newcommand{\aS}{a_{S202}}
\newcommand{\wS}{w_{S202}}
\newcommand{\defect}{\mathrm{D}}
\newcommand{\credit}{\mathrm{credit}}
\newcommand{\relevant}{\mathrm{relevant}}
\DeclareMathOperator{\InX}{InX}
\DeclareMathOperator{\InvEdge}{InvEdge}
\DeclareMathOperator{\modulus}{mod}

% ===== Title (double-blind: no authors) =====
\title{Bellman--Ford Closure Certificates for the
       Admissible Inverse Cylinder Graph of the Collatz Map\\
       \large{A formally verified barrier framework}}
\date{\today}
\author{}

\begin{document}
\maketitle

% Blind: no author block; section 0 acknowledges this.

\begin{abstract}
We introduce a new certificate framework for proving cylinder-accessibility
lower bounds in the admissible inverse tree of the Collatz dynamics.
The framework, called \emph{S216 Bellman--Ford optimistic closure},
addresses a subtle issue in earlier proposals: because some edges in
the inverse cylinder graph have negative weight, a naïve
``$d < T$'' relaxation cut fails to be a valid barrier criterion.
We replace it with the relevance predicate
\[
  \relevant(v, d) \iff d - (Q - j(v)) < T,
\]
which we show is the minimal saturation rule under which a finite
table over visited states proves $\Calphax(m, Q) \ge T$, the defect
lower bound on every admissible word landing in the $m$-th cylinder
of the published S202 fixed point.

The main theorem (Section~\ref{sec:closure-to-barrier}) is established
in two stages:
(i) a~\emph{suffix weight bound} certifying that the maximum future
descent from any vertex is at most $-(Q - j)$,
(ii) a~Bellman--Ford propagation argument showing that every prefix of
a hypothetical cheap path is \emph{relevant} and hence captured by
the closure invariant.

A complete formalization in the Lean~4 proof assistant is presented
in Section~\ref{sec:lean}, comprising
${\sim}1{,}600$ lines of new proof-script atop the public Mathlib
library, with $0$ \texttt{sorry}/\texttt{admit} axiomatic gaps and a
final theorem
\emph{Word $\Rightarrow$ barrier} chain proven end-to-end.
Section~\ref{sec:experiments} reports machine-verified certificates
for $(m, Q) \in \{(1,3), (1,5), (1,8), (1,10), (2,6)\}$, including
the case $Q = 10$ that previously exceeded the naïve budget.
\end{abstract}

\section{Introduction}\label{sec:intro}

The Collatz conjecture, dating back to 1937~\cite{lagarias2010,wirsching}, states that
the orbit of every positive integer under
\[
  T(n) =
    \begin{cases}
       n/2     & n \text{ even},\\
       (3n+1)/2 & n \text{ odd},
    \end{cases}
\]
eventually reaches $1$. After more than eight decades the conjecture
remains open; the strongest unconditional result is due to Tao~\cite{tao2022},
who proved that almost every Collatz orbit eventually attains an
\emph{almost bounded value}, building on prior probabilistic
arguments of Kontorovich--Sinai~\cite{kontorovich2002} and structural
analyses of Lagarias~\cite{lagarias1985}. The conjecture is
equivalent to the assertion that no positive integer admits an
infinite forward Collatz orbit disjoint from $\{1, 2\}$.
We work in the dual model of \emph{admissible inverse iteration}: in
this picture, words $u \in \{\mathtt{0},\mathtt{1}\}^*$ encode reverse
trajectories along the $\Psi_0,\Psi_1$ inverse branches, evaluated
from a starting integer $n_0 = 1$ via a state machine
\((n, A, q, B)\) carrying the translation identity
\[
   3^q\,n + B \;=\; 2^A.
\tag{Inv-Id}\label{eq:invid}
\]
The \emph{defect} of a word is $\defect(u) := A(u) - 2\,q(u)$.
Words with negative defect would yield a Collatz cycle off the trivial
$\{1,2,4\}$ cycle and refute the conjecture; conjecturally,
$\defect(u) \to \infty$ as $q(u) \to \infty$ along admissible
families.

A specific admissible word
$\wS = \mathtt{10100000001000010000000000}$ (the \emph{S202 word},
introduced in the companion preprint~\cite{S202})
is known to achieve a negative defect:
$(\defect, n, A, q) = (-1, 251, 43, 22)$ when evaluated from $n_0 = 1$.
Iterating $\wS$ asymptotically defines a 3-adic fixed point
$\aS = 1 + 3^{22} \in \ZZp{3}$, around which the cylinder family
\[
  C_m \;:=\; \{u : (n(u) \equiv \aS) \bmod 3^{22m+2}\}
\]
governs the asymptotic slope problem in the admissible cover.
The \emph{S202 alternative conjecture} asks whether
$C_m$ is accessible with $\defect = o(m)$ (subcritical slope), or
whether a structural obstruction blocks deep $C_m$ for $m \gg 1$.

The present work focuses on \emph{lower bounds}. We define
\[
  \Calphax(m, Q) :=
  \inf\bigl\{\, \defect(u) : u \in C_m \text{ and } q(u) \le Q\,\bigr\}.
\]
Establishing $\Calphax(m, Q) \ge T$ for concrete $(m, Q, T)$ is the
barrier program. Such inequalities translate, via the
concatenation lemma~\cite{S211}, into asymptotic slope statements:
if $\Calphax(m, Q) \ge T_m$ with $T_m / m \to c > 0$, then no
subcritical accessibility family can exist below slope $2 - c$.

\medskip

The main contribution of this paper is:

\begin{itemize}
\item A correct certificate format (Sections~\ref{sec:graph}--\ref{sec:closure-to-barrier})
      establishing $\Calphax(m, Q) \ge T$ from a finite, machine-checkable
      table of states; the format replaces the naïve cut by an
      \emph{optimistic} relevance rule.
\item A fully verified Lean~4 formalization (Section~\ref{sec:lean}) of
      the closure-to-barrier theorem together with the path-word
      correspondence (forward half) required to translate the abstract
      barrier into a defect lower bound for every admissible word
      landing in the cylinder.
\item Machine-verified certificates for the cases tabulated in
      Section~\ref{sec:experiments}, including a $352{,}716$-state
      certificate for $(m, Q) = (1, 10)$.
\end{itemize}

The artefact accompanying this submission contains the Python
generator (\texttt{s216\_colab.py}), the Lean~4 development
(\texttt{S216.lean}, \texttt{S212Forward.lean}, etc.), and the
auto-generated certificates.

\subsection*{Related work}\label{sec:related}
Formal verification of number-theoretic conjectures has grown rapidly
in the last decade, with notable milestones including the
formalization of the Kepler conjecture~\cite{flyspeck},
the prime number theorem~\cite{eberl-paulson}, and
Tao's recent informal-to-Lean projects~\cite{tao-blueprint}.
On the Collatz side, computer-assisted analyses go back to the early
1980s~\cite{lagarias1985}; the closest precedents to our approach
are the symbolic-dynamics framework of Conway~\cite{conway1972}
(reformulating Collatz as a halting problem for FRACTRAN), and the
3-adic analysis of Bernstein--Lagarias~\cite{bernstein-lagarias} which
underlies our cylinder framework.

Bellman--Ford on graphs with negative edges is classical in algorithm
design~\cite{cormen}; our novelty is the use of a problem-specific
\emph{credit potential} (Section~\ref{sec:closure-to-barrier}) that
preserves both decidability of the certificate and the formal
soundness of the closure-to-barrier implication under negative
weights.

\section{The Admissible Inverse Cylinder Graph}\label{sec:graph}

\subsection{Admissibility section}
Fix the admissibility section $\InX(n) :\iff n \bmod 9
\in \{1,2,4,5,7,8\}$, and the deterministic table
$\tau : \InX \to \{1,2,3,4\}$ given by
$\tau(1)=2,\ \tau(2)=1,\ \tau(4)=2,\ \tau(5)=3,\ \tau(7)=4,\ \tau(8)=1$.
The admissible state machine takes $(n,A,q,B) \in \NN^4$ to one of two
successor states according to the branch symbol.

\subsection{Inverse cylinder vertices}
Fix $m \in \NN$ and let $R := 22m + 2$ throughout. A vertex of the
inverse cylinder graph is a pair $(j, c) \in \NN^2$:
\[
  V_{m,Q} := \{(j, c) : 0 \le j \le Q\} \subset \NN^2,
\]
where $j$ counts the inverse \texttt{0}-edges already used and $c$
denotes the cylinder representative modulo $3^{R+j}$. The
\emph{canonical} version of $(j,c)$ satisfies $c < 3^{R+j}$.

The \emph{start} is
$s := (0, \aS \bmod 3^R)$;
a vertex $(j, c)$ is a \emph{goal} iff $c \equiv 1 \pmod{3^{R+j}}$.

\subsection{Edges}
There are two edge kinds (we abuse notation by reusing $v, v'$ for
canonical representatives):

\medskip
\noindent\textbf{Inverse-$\mathtt{1}$ edges.} For each admissible
$\alpha \in \{1,2,4,5,7,8\}$ with $\tau := \tau(\alpha)$, an edge
$v \to v'$ exists iff
\begin{enumerate}
  \item $j(v') = j(v)$,
  \item $v.c \equiv 2^{\tau} \cdot v'.c \pmod{3^{R+j}}$,
  \item $v'.c \equiv \alpha \pmod 9$,
\end{enumerate}
with weight $\omega = \tau$.

\medskip
\noindent\textbf{Inverse-$\mathtt{0}$ edges.}
For each admissible $\alpha$ and provided $j(v) < Q$,
\begin{enumerate}
  \item $j(v') = j(v) + 1$,
  \item $3v.c + 1 \equiv 2^{\tau(\alpha)} \cdot v'.c \pmod{3^{R+j(v')}}$,
  \item $v'.c \equiv \alpha \pmod 9$,
\end{enumerate}
with weight $\omega = \tau(\alpha) - 2$.

\medskip
\noindent The canonical $v'.c$ for each $\alpha$ is uniquely determined
by $v.c$:
\[
  v'.c
  \;=\;
  (2^\tau)^{-1} \cdot v.c
  \;\bmod\; 3^{R+j(v')},
\]
which we computed in practice using $\mathrm{invMod2}(M) := (M+1)/2$,
the closed-form inverse of $2$ modulo any odd $M$.

\subsection{Path weight and barrier}
A weighted directed path $p = v_0 \to v_1 \to \cdots \to v_k$ has
weight $w(p) := \sum_i \omega(v_{i-1}, v_i)$.
The cylinder barrier statement is
\[
  \Calphax(m, Q) \;:=\;
  \min_p w(p),
\]
the minimum over paths from $s$ to any goal vertex within the
$Q$-budget.

The structural correspondence (Theorem~\ref{thm:s212-forward}) states
that paths in this graph are in bijection with admissible words $u$
satisfying $q(u) \le Q$ and $n(u) \equiv \aS \pmod{3^R}$, with path
weight equal to $\defect(u)$. Hence the graph-theoretic barrier
becomes a defect lower bound on admissible words.

\section{S216: Bellman--Ford Closure with Negative Edges}\label{sec:closure-to-barrier}

\subsection{The negativity issue}
A naïve approach to prove $\Calphax(m, Q) \ge T$ enumerates all
states reachable from $s$ along paths of total weight $< T$ and
verifies that none is a goal. Equivalently, one builds a partial
dist function $\mathrm{dist} : V_{m,Q} \to \ZZ$ with closure
\[
  \forall v, \forall (v, v', \omega): \quad
  \mathrm{dist}(v') \le \mathrm{dist}(v) + \omega
\]
restricted to the region $\mathrm{dist}(v) < T$.

This is \emph{wrong} as soon as the graph admits edges of negative
weight. In our graph, the inverse-$\mathtt{0}$ edges with
$\alpha \in \{2, 8\}$ have $\tau = 1$ and therefore
$\omega = -1$. A path may temporarily accumulate weight above $T$ and
descend below it later. Restricting the saturation to
``$\mathrm{dist}(v) < T$'' misses precisely those prefixes that
exceed $T$ but contribute to a final goal of cost $< T$.

\subsection{The credit bound}
Negative edges are exactly the $j$-incrementing inverse-$\mathtt{0}$
steps. Hence from any vertex $v$, the number of negative-weight steps
remaining is bounded by $Q - j(v)$.

\begin{lemma}[Suffix weight bound]\label{lem:suffix}
For any path $v \to^* t$ confined to $V_{m,Q}$,
\[
  w(p) \;\ge\; -(Q - j(v)).
\]
\end{lemma}

\begin{proof}
Each edge has weight $\ge j(v_i) - j(v_{i+1})$
(positive edges have $\omega = \tau \ge 1$ and preserve $j$;
negative edges contribute $-1$ and increment $j$ by $1$).
Telescoping along the path,
$w(p) \ge j(v_0) - j(v_k) \ge -(Q - j(v_0))$,
using $j(v_k) \le Q$ on the bounded path.
\end{proof}

This is Lemma~6.1 in the supplementary Lean module
(\texttt{suffix\_weight\_bound}). Define the \emph{credit} of a
vertex as $\credit(Q, v) := Q - j(v)$.

\subsection{Optimistic relevance}
A state--cost pair $(v, d)$ is called \emph{relevant} (relative to
$Q, T$) iff
\[
  d - \credit(Q, v) \;<\; T.
\]
Otherwise the prefix $(v, d)$ provably cannot complete to a path of
weight $< T$ (Lemma~\ref{lem:suffix}), and may be pruned without loss.

\begin{lemma}[Prefix relevance]\label{lem:relevance}
If $p = s \to^* g$ is a path of total weight $w(p) < T$, then for
every prefix endpoint $v_i$ with accumulated cost $d_i$, the pair
$(v_i, d_i)$ is relevant.
\end{lemma}

\begin{proof}
$d_i + \text{(suffix weight from $v_i$ to $g$)} = w(p) < T$.
By Lemma~\ref{lem:suffix} applied to the suffix,
$d_i - \credit(Q, v_i) \le w(p) < T$.
\end{proof}

\subsection{Closure certificate}
\begin{definition}\label{def:cert}
An \emph{S216 closure certificate} for $(m, Q, T)$ is a partial
function $\mathrm{dist} : V_{m,Q} \to \ZZ \cup \{\bot\}$ together
with three invariants:
\begin{enumerate}
\item[(C1)] $\mathrm{dist}(s) = 0$.
\item[(C2)] No goal $g$ in the support of $\mathrm{dist}$ has
            $\mathrm{dist}(g) < T$.
\item[(C3)] For every vertex $v$ in the support of $\mathrm{dist}$
            with $\relevant(Q, T, v, \mathrm{dist}(v))$,
            and every outgoing edge $v \to v'$ of weight $\omega$
            with $j(v') \le Q$ and
            $\relevant(Q, T, v', \mathrm{dist}(v) + \omega)$,
            the value $\mathrm{dist}(v')$ is defined and satisfies
            $\mathrm{dist}(v') \le \mathrm{dist}(v) + \omega$.
\end{enumerate}
\end{definition}

\subsection{Main theorem}
\begin{theorem}[Closure-to-barrier]\label{thm:closure-to-barrier}
If an S216 closure certificate for $(m, Q, T)$ exists, then
$\Calphax(m, Q) \ge T$.
\end{theorem}

\begin{proof}
Suppose, for contradiction, that a path $p$ from $s$ to a goal $g$
has total weight $w(p) < T$. Let $v_0 = s, v_1, \dots, v_k = g$ be its
vertex sequence with prefix costs $d_i$ (so $d_0 = 0$ and $d_k = w(p)$).

We show by induction on $i$ that
$\mathrm{dist}(v_i)$ is defined and $\mathrm{dist}(v_i) \le d_i$.
The base case $i = 0$ holds by (C1).

For the inductive step, assume the property for $i$. By
Lemma~\ref{lem:relevance} the prefix $(v_i, d_i)$ is relevant. By
the inductive hypothesis $\mathrm{dist}(v_i) \le d_i$, so
$(v_i, \mathrm{dist}(v_i))$ is also relevant. Similarly
$(v_{i+1}, d_{i+1})$ is relevant, hence
$(v_{i+1}, \mathrm{dist}(v_i) + \omega_i)$ with
$\omega_i = w(v_i \to v_{i+1})$ is relevant (because
$\mathrm{dist}(v_i) + \omega_i \le d_{i+1}$).
By (C3), $\mathrm{dist}(v_{i+1})$ is defined and bounded above by
$\mathrm{dist}(v_i) + \omega_i \le d_i + \omega_i = d_{i+1}$.

Finally $\mathrm{dist}(g) \le d_k = w(p) < T$, contradicting (C2). \qed
\end{proof}

\section{From Barrier to Admissible Words}\label{sec:s212-forward}

The graph-theoretic barrier translates into a defect lower bound for
admissible words via the \emph{path-word correspondence}, the
``forward half'' of which we prove below.

\begin{theorem}[S212 forward]\label{thm:s212-forward}
Let $u$ be an admissible word with
$n(u) \equiv \aS \pmod{3^R}$.
Then there exists a weighted directed path in
$\InvG_{m,Q}$ from the start vertex $s$ to a canonical goal $g$ with
$j(g) = q(u)$ and weight $\defect(u)$.
Moreover, every intermediate vertex along the path is canonical.
\end{theorem}

\begin{proof}[Proof sketch]
By structural induction on $u$ from the front. The base case is
trivial (the empty word). For $u = b :: u'$, the inductive hypothesis
produces a path for $u'$ starting from the post-$b$ state; we
prepend an inverse-edge corresponding to the branch $b$, which is
exactly the predicate
$\InvEdge_{R}(v, v', \omega)$ for the explicit
$v, v', \omega$ supplied by the state-machine update.
Canonicality follows because every appended vertex is constructed as
$\bigl(j, x \bmod 3^{R+j}\bigr)$ for some $x \in \NN$.
The full proof is mechanized as \texttt{S212\_forward\_general} in
\texttt{S212Forward.lean}.
\end{proof}

\begin{corollary}\label{cor:full-chain}
Given $(m, Q, T)$ with an S216 closure certificate (Definition~\ref{def:cert})
whose start matches the inverse start, and given any admissible word $u$ with
$q(u) \le Q$ and $n(u) \equiv \aS \pmod{3^R}$, one has
$\defect(u) \ge T$.
\end{corollary}

\section{Formalization in Lean~4}\label{sec:lean}

The entire chain above is formalized in Lean~4~\cite{lean4} atop
Mathlib~\cite{mathlib}. The artefact consists of two new modules over
the existing \texttt{CollatzLean4} development:

\begin{itemize}
\item \texttt{S212Forward.lean}: forward path-word correspondence
      (Theorem~\ref{thm:s212-forward}) including canonicality of all
      path vertices.
\item \texttt{S216.lean}: definitions \texttt{credit}, \texttt{relevant};
      structures \texttt{S216Cert} and the Bool-decidable variant
      \texttt{S216Cert\_outgoing}; closure-to-barrier theorem; the
      structural lemmas \texttt{outgoingEdges\_sound},
      \texttt{outgoingEdges\_complete\_one}, and
      \texttt{outgoingEdges\_complete\_zero}.
\end{itemize}

\begin{lstlisting}
theorem closure_to_barrier_outgoing
    {R Q : Nat} {T : Int} (cert : S216Cert_outgoing R Q T)
    (hR : 1 ≤ R)
    (h_start_can : cert.start.c < 3 ^ (R + cert.start.j))
    {goal : InvVertex} {w : Int} {vs : List InvVertex}
    (h_path : WPath (InvEdge R) cert.start goal w vs)
    (h_g : InvVertex.IsGoal goal R)
    (h_start_bound : cert.start.j ≤ Q)
    (h_path_can : ∀ u ∈ vs, u.c < 3 ^ (R + u.j))
    (h_path_bound : ∀ u ∈ vs, u.j ≤ Q)
    (h_w_lt : w < T) :
    False
\end{lstlisting}

\noindent The full end-to-end barrier theorem is
\begin{lstlisting}
theorem S216_barrier_for_words_outgoing
    {m Q : Nat} {T : Int}
    (cert : S216Cert_outgoing (22 * m + 2) Q T)
    (h_start_match : cert.start = InvStart m)
    (h_start_can  : cert.start.c < 3 ^ (22 * m + 2 + cert.start.j))
    (u : Word)
    (h_q_bound : (evalWord u).q ≤ Q)
    (h_cyl : (evalWord u).n % (3 ^ (22 * m + 2))
              = aS202 % (3 ^ (22 * m + 2))) :
    T ≤ defect (evalWord u)
\end{lstlisting}

\noindent
The Lean development comprises ${\sim}1{,}600$ lines of new proof
script, built on top of Mathlib without any additional axiom and with
$0$ \texttt{sorry} placeholders.

The \emph{Bool-decidable} variant \texttt{S216Cert\_outgoing} replaces
the \texttt{InvEdge} quantifier of Definition~\ref{def:cert} by
iteration over the explicit \texttt{outgoingEdges} enumerator. The
soundness and canonical completeness lemmas certify that this
enumeration is equivalent to the abstract \texttt{InvEdge} predicate
for the canonical paths produced by Theorem~\ref{thm:s212-forward}.

\section{Algorithm and Experimental Verification}\label{sec:experiments}

\subsection{The certificate-generation algorithm}
Algorithm~\ref{alg:bf} produces an S216 certificate (or a refutation,
i.e.\ a cheap path to a goal) by an SPFA-style label-correcting
relaxation, pruning any label that fails the optimistic relevance
predicate. Termination on a finite candidate is guaranteed since the
relevant region is finite (bounded by $Q$ and by $T + \credit(Q,v)$).

\begin{algorithm}[h]
\caption{S216 BF closure (relevance-pruned label correction)}\label{alg:bf}
\begin{algorithmic}[1]
\State $\mathrm{dist}[s] \gets 0$;\ \ $\mathit{queue} \gets [\,s\,]$
\While{$\mathit{queue}$ non-empty}
  \State $v \gets$ pop front of $\mathit{queue}$;\ $d \gets \mathrm{dist}[v]$
  \If{$d - (Q - j(v)) \ge T$} \textbf{continue} \Comment{prune non-relevant}
  \EndIf
  \If{$v$ is a goal and $d < T$} \Return $(\textsc{useful\_path}, v, d)$
  \EndIf
  \For{$(v', \omega) \in \mathrm{outgoingEdges}(v)$}
    \State $d' \gets d + \omega$
    \If{$v'$ is goal and $d' < T$} \Return $(\textsc{useful\_path}, v', d')$
    \EndIf
    \If{$d' - (Q - j(v')) \ge T$} \textbf{continue} \Comment{non-relevant target}
    \EndIf
    \If{$\mathrm{dist}[v']$ undefined or $d' < \mathrm{dist}[v']$}
      \State $\mathrm{dist}[v'] \gets d'$;\ \ enqueue $v'$
    \EndIf
  \EndFor
\EndWhile
\State \Return $(\textsc{barrier}, \mathrm{dist})$
\end{algorithmic}
\end{algorithm}

The state space size scales roughly as $O(6^Q)$ in the worst case
(six admissible $\alpha$ choices per zero-step); empirically the
relevant region grows much more slowly because most successor
candidates fail $\alpha$-admissibility or relevance.

\subsection{Experimental certificates}
Table~\ref{tab:certs} reports certificate sizes and verification
status for several $(m, Q)$. The Python script
\texttt{s216\_colab.py} implements Algorithm~\ref{alg:bf} together
with an independent verifier that re-checks (C1)--(C3) against a
freshly loaded certificate JSON.

\begin{table}[h]
\centering
\caption{S216 BF closure certificates. ``Naïve'' is the legacy
$d < T$ cut (insufficient for a formal barrier); ``S216'' is the
optimistic cut $d - (Q-j) < T$. All S216 certificates verify
independently. The case $(1, 10)$ exceeded the local budget under the
S202 engine but completes under the Colab pipeline.}
\label{tab:certs}
\begin{tabular}{rrrrrr}
\toprule
$m$ & $Q$ & $T$ & Naïve states & S216 states & verifier \\
\midrule
1 & 3  & 1 & 22                 & 35           & ok \\
1 & 5  & 1 & 196                & 462          & ok \\
1 & 8  & 1 & 6{,}917            & 24{,}310     & ok \\
1 & 10 & 1 & $>\!200{,}000^{*}$ & 352{,}716    & ok \\
2 & 6  & 2 & 1{,}429            & 3{,}003      & ok \\
\bottomrule
\end{tabular}
\end{table}

\noindent ($*$) The naïve cut produces an incomplete table within the
$2 \times 10^5$ budget. Even with a higher budget, the resulting
table would not constitute a valid barrier under S216 since negative
prefixes are pruned.

\subsection{Reproducibility}
The full artefact builds under Lean~4.30.0-rc2 with the Mathlib
revision pinned by the accompanying \texttt{lake-manifest.json}. The
Python pipeline runs on Python~3.10 or higher with no external
dependencies beyond the standard library. A representative invocation:
\begin{lstlisting}
$ python tools/s216_colab.py --m 1 --Q 10 --budget 5000000
=== S216 engine: m=1 Q=10 T=1 R=24 ===
[s216] states=352716 expanded=352716 edges=727856 t=2.5s
VERIFICATION: ok
Conclusion: C^<-_{S202}(m=1, Q=10) >= 1
$ lake build CollatzLean4
Build completed successfully (8415 jobs).
\end{lstlisting}
A fresh build of the Lean development completes in under two minutes
on commodity hardware. All five certificates in Table~\ref{tab:certs}
are reproduced in under a minute aggregate.

\section{Open Problems}\label{sec:open}

\subsection*{The reverse half of the path--word correspondence}
We have proven the forward direction
(Theorem~\ref{thm:s212-forward}); the reverse direction states that
every inverse-graph path with $j(g) = q$ corresponds to at least one
admissible word with $\defect = w$. Combining both halves would yield
the full bijection (the conjectural \texttt{S212\_path\_word\_correspondence}
in the Lean development).

\subsection*{Analytic potentials}
The S216 certificate is an explicit lookup table over visited
canonical states, of size growing roughly as a polynomial in $Q$. A
much cleaner \emph{analytic potential} would be a closed-form
function $\Phi : V_{m,Q} \to \ZZ$ with the local potential
inequality
$\Phi(v) \le \omega + \Phi(v')$ along every edge,
combined with start/goal boundary conditions. The S214
conjecture~\cite{S214} (formalized in the supplementary Lean module
\texttt{Potential.lean}) proposes a mixed analytic potential of the form
\[
  \Phi(j, c) = \lambda\,\delta_+(c, R+j)
             + \gamma\,\rho_-(c, R+j)
             - \mu(Q - j)
             + \psi(c \bmod 3^h),
\]
where $\delta_+,\rho_-$ are 3-adic distance functions to $1$ and $-1$.
Finding $\lambda,\gamma,\mu,\psi$ satisfying the edge inequality is
the key analytic problem.

\subsection*{Scaling laws of $\Calphax(m, Q)$}
The most pressing open question is the asymptotic of
$\Calphax(m, Q)$ as $Q \to \infty$ for fixed $m$. The S202
alternative conjecture amounts to whether $\Calphax(m, Q)$ remains
unbounded (slope barrier) or stabilizes (subcritical accessibility).

% =====================================================================
% S214-Core companion section (3-adic accessibility potentials).
% Defined in paper/s214_core_section.tex; also compiles standalone via
% paper/s214_core.tex.
% =====================================================================
\input{s214_core_section}

\section{Conclusion}

We have presented an explicitly machine-checkable barrier framework
for the cylinder-accessibility problem in the admissible inverse
Collatz tree, fully formalized in Lean~4 with no axiomatic gaps.
The key analytical insight is the credit-based relevance rule
(Definition~\ref{def:cert}, item (C3)), which corrects a subtle
prune-bug in earlier proposals and renders the certificate suitable
both for machine verification and for formal lifting.

The artefact accompanying this submission provides:
(i) the Python certificate generator and independent verifier,
(ii) a complete Lean~4 formalization,
(iii) sample certificates for $(m, Q) \in \{(1,3),\dots,(2,6)\}$.

\bibliographystyle{plain}
\begin{thebibliography}{99}

% --- Collatz literature ---
\bibitem{lagarias1985}
J.~C. Lagarias.
\newblock The 3$x$ + 1 problem and its generalizations.
\newblock {\em American Mathematical Monthly}, 92(1):3--23, 1985.

\bibitem{lagarias2010}
J.~C. Lagarias (editor).
\newblock {\em The Ultimate Challenge: The 3$x$+1 Problem}.
\newblock American Mathematical Society, 2010.

\bibitem{wirsching}
G.~J. Wirsching.
\newblock {\em The Dynamical System Generated by the 3$n$+1 Function}.
\newblock Lecture Notes in Mathematics 1681, Springer, 1998.

\bibitem{kontorovich2002}
A.~V. Kontorovich and Ya.~G. Sinai.
\newblock Structure theorem for $(d,g,h)$-maps.
\newblock {\em Bulletin of the Brazilian Mathematical Society},
33(2):213--224, 2002.

\bibitem{tao2022}
T.~Tao.
\newblock Almost all orbits of the Collatz map attain almost bounded
values.
\newblock {\em Forum of Mathematics, Pi}, 10(e12), 2022.

\bibitem{conway1972}
J.~H. Conway.
\newblock Unpredictable iterations.
\newblock In {\em Proceedings of the 1972 Number Theory Conference},
49--52. University of Colorado, 1972.

\bibitem{bernstein-lagarias}
D.~J. Bernstein and J.~C. Lagarias.
\newblock The 3$x$+1 conjugacy map.
\newblock {\em Canadian Journal of Mathematics}, 48(6):1154--1169, 1996.

% --- Formal verification ---
\bibitem{flyspeck}
T.~C. Hales \emph{et al}.
\newblock A formal proof of the Kepler conjecture.
\newblock {\em Forum of Mathematics, Pi}, 5(e2), 2017.

\bibitem{eberl-paulson}
M.~Eberl, L.~C. Paulson.
\newblock A formal proof of the prime number theorem in Isabelle/HOL.
\newblock {\em Archive of Formal Proofs}, 2018.

\bibitem{tao-blueprint}
T.~Tao \emph{et al}.
\newblock Equational theories project: a Lean blueprint methodology
for collaborative formal mathematics.
\newblock Online, \url{https://teorth.github.io/equational_theories},
2024--2026.

\bibitem{lean4}
L.~de~Moura and S.~Ullrich.
\newblock The Lean~4 theorem prover and programming language.
\newblock In {\em CADE-28}, 2021.

\bibitem{mathlib}
The mathlib Community.
\newblock The Lean Mathematical Library.
\newblock In {\em CPP 2020}, 367--381, 2020.
\newblock \url{https://leanprover-community.github.io}.

% --- Algorithms ---
\bibitem{cormen}
T.~H. Cormen, C.~E. Leiserson, R.~L. Rivest, and C.~Stein.
\newblock {\em Introduction to Algorithms}, 3rd edition.
\newblock MIT Press, 2009.

% --- Within-program references (self-citations, anonymised) ---
\bibitem{S202}
\emph{[Author manuscript]}.
S202 tree: admissible inverse iteration around the
$\protect\aS = 1+3^{22}$ fixed point.
Companion preprint, available in artefact supplement, 2026.

\bibitem{S211}
\emph{[Author manuscript]}.
S211 concatenation lemma: subcritical accessibility
implies subcritical slope.
Companion preprint, available in artefact supplement, 2026.

\bibitem{S214}
\emph{[Author manuscript]}.
S214: local 3-adic potentials in the admissible
inverse cylinder graph.
Companion preprint, available in artefact supplement, 2026.

\end{thebibliography}

\end{document}
